---
title: "STATS 506_HW1"
author: "Yixuan Jia"
date: "2023-09-09"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Github Link: https://github.com/CathyJia128/repo_1

Problem 1:
```{r}
data <- read.table('wine.data', header = FALSE, sep = ",")

col_names <- c(
  "Class", "Alcohol", "Malic acid", "Ash",
  "Alcalinity of ash", "Magnesium", "Total phenols", "Flavanoids",
  "Nonflavanoid phenols", "Proanthocyanins", "Color intensity", "Hue",
  "OD280/OD315 of diluted wines", "Proline"
)
colnames(data) <- col_names
```

(c)：  
(1) The wine with the highest alcohol content belongs to which class?
```{r}
highest <- data$Class[which.max(data$Alcohol)]

cat("The wine has the highest alcohol belongs to Class", highest)
```

(2) The wine with the lowest alcohol content belongs to which class?
```{r}
lowest <- data$Class[which.min(data$Alcohol)]

cat("The wine has the lowest alcohol belongs to Class", lowest)
```

(3) German beers have, on average, 114 mg/l of magnesium. How many of these wines have higher levels of magnesium than that? (You may assume that the magnesium reported in the data is in the same units.)
```{r}
cat("The number of wines have higher levels of magnesium than German beers is:", sum(data$Magnesium > 114))
```

(4) Within each class, how many wines have higher levels of magnesium than average German beer?
```{r}
c1 <- aggregate(Magnesium ~ Class, data, function(x) sum(x > 114))
c1
```

(d):  
```{r}
overall <- c(0, colMeans(data[,-1])) 
class <- aggregate(. ~ Class, data = data, FUN = mean)
t1 <-rbind('Overall Average' = overall, 'Class Average' = class)
t1
```
(e):  
```{r}
class_pairs <- combn(unique(data$Class), 2)

for (i in 1:ncol(class_pairs)) {
  class1 <- class_pairs[1, i]
  class2 <- class_pairs[2, i]
  ash1 <- data$Ash[data$Class == class1]
  ash2 <- data$Ash[data$Class == class2]
  result <- t.test(ash1, ash2)
  
  cat("Class", class1, "vs. Class", class2, ":\n")
  cat("p-value:", result$p.value, "\n")
  cat("\n")
}
```
According to the results of the test, we can see that the p-value for class 1 and class 2 is very small (less than 0.05), so we can conclude that there are significant difference between the level of Ash in class 1 and 2.  
Same situation for class 2 and 3, the p-value is less than 0.05, so the level of Ash differs across Class 2 and 3.  
However, the p-value for class 1 and 3 is 0.64, which is greater than 0.05, so there is no significant difference in Ash between these two classes.

Problem 2:
(a):  
```{r}
isPerfectPower <- function(x, power) {
  root <- round(x^(1/power))
  if (root^power == x) {
    return(list(isPerfect = TRUE, root = root))
  }
  else {
    return(list(isPerfect = FALSE, root = NA))
  }
}
```

(b):  
```{r}
findRootPower <- function(x) {
  power <- 2  # Start with power = 2
  while (power <= 1000) {
    result <- isPerfectPower(x, power)
    if (result$isPerfect) {
      return(paste(x, "=", result$root, "^", power))
    }
    power <- power + 1  # Increment the power
  }
  return(paste(x, "is not a perfect power within the specified range."))
}

print(findRootPower(27))
print(findRootPower(13060694016))
print(findRootPower(7776))
print(findRootPower(170859375))
print(findRootPower(58247422))
print(findRootPower(94143178827))
```

Problem 3:  
a)
```{r}
# Function to determine the name of a poker hand
get_poker_hand_name <- function(suits, ranks) {
  # Check for a flush (all cards of the same suit)
  if (length(unique(suits)) == 1) {
    flush <- TRUE
    suit <- unique(suits)
  } else {
    flush <- FALSE
    suit <- NULL
  }
  
  # Count the frequency of each rank
  rank_counts <- table(ranks)
  
  # Check for four of a kind (four cards with the same rank)
  if (any(rank_counts == 4)) {
    return("Four of a Kind")
  }
  
  # Check for a full house (three cards with one rank and two cards with another rank)
  if (any(rank_counts == 3) && any(rank_counts == 2)) {
    return("Full House")
  }
  
  # Check for a flush
  if (flush) {
    return("Flush")
  }
  
  # Check for three of a kind (three cards with the same rank)
  if (any(rank_counts == 3)) {
    return("Three of a Kind")
  }
  
  # Check for two pair (two cards with one rank and two cards with another rank)
  if (sum(rank_counts == 2) == 4) {
    return("Two Pair")
  }
  
  # Check for one pair (two cards with the same rank)
  if (any(rank_counts == 2)) {
    return("One Pair")
  }
  
  # If none of the above conditions are met, it's a high card hand
  return("High Card")
}

# Function to simulate a round of poker for a specified number of players
deal_poker_round <- function(num_players) {
  if (num_players < 2 || num_players > 10) {
    cat("Number of players must be between 2 and 10.")
    return(NULL)
  }
  
  # Define the deck of cards
  suits <- c("Hearts", "Diamonds", "Clubs", "Spades")
  ranks <- c(2:10, "Jack", "Queen", "King", "Ace")
  deck <- expand.grid(rank = ranks, suit = suits)

  # Shuffle the deck
  shuffled_deck <- deck[sample(nrow(deck)), ]
  
  # Deal cards to each player without replacement
  player_hands <- vector("list", num_players)
  for (i in 1:num_players) {
    # Randomly choose 5 unique cards from the shuffled deck
    player_indices <- sample(1:nrow(shuffled_deck), size = 5, replace = FALSE)
    player_hand <- shuffled_deck[player_indices, ]
    
    # Remove the dealt cards from the shuffled deck
    shuffled_deck <- shuffled_deck[-player_indices, ]
    
    cat("Player", i, "Hand:")
    print(player_hand)
    hand_name <- get_poker_hand_name(player_hand$suit, player_hand$rank)
    cat("Hand Name:", hand_name, "\n\n")
    
    player_hands[[i]] <- player_hand
  }
}

# Example usage:
deal_poker_round(4)  # Simulate a round of poker with 4 players
```

b) First, it gave me an error, but I told ChatGPT what error it generated and ChatGPT fixed it. 
```{r}
deal_poker_round(3)  # Simulate a round of poker with 3 players
deal_poker_round(4)  # Simulate a round of poker with 4 players
deal_poker_round(5)  # Simulate a round of poker with 5 players
```

c)
```{r}
# Function to determine the name of a poker hand
get_poker_hand_name <- function(suits, ranks) { #Defines a function "get_poker_hand_name", with two inputs: suits and ranks
  
  # Check for a flush (all cards of the same suit)
  if (length(unique(suits)) == 1) {  #it checks whether all cards have a same suit
    flush <- TRUE #if so, it returns "flush" 
    suit <- unique(suits) #record the suit
  } else {
    flush <- FALSE #otherwise
    suit <- NULL
  }
  
  # Counts how many cards of each rank are in the hand
  rank_counts <- table(ranks) 
  
  # Check for four of a kind (four cards with the same rank)
  if (any(rank_counts == 4)) {
    return("Four of a Kind")
  }
  
  # Check for a full house (three cards with one rank and two cards with another rank)
  if (any(rank_counts == 3) && any(rank_counts == 2)) {
    return("Full House")
  }
  
  # Check for a flush
  if (flush) {
    return("Flush")
  }
  
  # Check for three of a kind (three cards with the same rank)
  if (any(rank_counts == 3)) {
    return("Three of a Kind")
  }
  
  # Check for two pair (two cards with one rank and two cards with another rank)
  if (sum(rank_counts == 2) == 4) {
    return("Two Pair")
  }
  
  # Check for one pair (two cards with the same rank)
  if (any(rank_counts == 2)) {
    return("One Pair")
  }
  
  # If none of the above conditions are met, it's a high card hand
  return("High Card")
}

# Function to simulate a round of poker for a specified number of players
deal_poker_round <- function(num_players) {
  if (num_players < 2 || num_players > 10) { # if num_player is not between 2 and 10
    cat("Number of players must be between 2 and 10.") # print out the comment
    return(NULL) # and return null
  }
  
  # Define the deck of cards
  suits <- c("Hearts", "Diamonds", "Clubs", "Spades")
  ranks <- c(2:10, "Jack", "Queen", "King", "Ace")
  deck <- expand.grid(rank = ranks, suit = suits) #creates a data frame that represents all the possible combinations of ranks and suits 
  
  if (num_players * 5 > nrow(deck)) {
    cat("Not enough cards in the deck to deal to all players.")
    return(NULL)
  }
  
  # Shuffle the deck
  #set.seed(42) # For reproducibility
  shuffled_deck <- deck[sample(nrow(deck)), ] #used "sample" function to resample the rows in the data frame "deck", so that it can generate a random set of cards

  # Deal cards to each player without replacement
  player_hands <- vector("list", num_players) #This is for initialization, the "player_hands" is an empty list, used to store the future results.
  for (i in 1:num_players) {
    # Randomly choose 5 unique cards from the shuffled deck
    player_indices <- sample(1:nrow(shuffled_deck), size = 5, replace = FALSE) 
    player_hand <- shuffled_deck[player_indices, ]
    
    # Remove the dealt cards from the shuffled deck (this line of code ensures no duplicate)
    shuffled_deck <- shuffled_deck[-player_indices, ]
    
    cat("Player", i, "Hand:")
    print(player_hand)
    hand_name <- get_poker_hand_name(player_hand$suit, player_hand$rank)
    cat("Hand Name:", hand_name, "\n\n")  #print out the results
    
    player_hands[[i]] <- player_hand
  }
}

# Example usage:
deal_poker_round(4)  # Simulate a round of poker with 4 players
```

 
d)
```{r}
set.seed(42)
# Function to determine the name of a poker hand
get_poker_hand_name <- function(suits, ranks) {
  # Check for a flush (all cards of the same suit)
  if (length(unique(suits)) == 1) {
    flush <- TRUE
    suit <- unique(suits)
  } else {
    flush <- FALSE
    suit <- NULL
  }
  
  # Count the frequency of each rank
  rank_counts <- table(ranks)
  
  # Check for four of a kind (four cards with the same rank)
  if (any(rank_counts == 4)) {
    return("Four of a Kind")
  }
  
  # Check for a full house (three cards with one rank and two cards with another rank)
  if (any(rank_counts == 3) && any(rank_counts == 2)) {
    return("Full House")
  }
  
  # Check for a flush
  if (flush) {
    return("Flush")
  }
  
  # Check for three of a kind (three cards with the same rank)
  if (any(rank_counts == 3)) {
    return("Three of a Kind")
  }
  
  # Check for two pair (two cards with one rank and two cards with another rank)
  if (sum(rank_counts == 2) == 4) {
    return("Two Pair")
  }
  
  # Check for one pair (two cards with the same rank)
  if (any(rank_counts == 2)) {
    return("One Pair")
  }
  
  # If none of the above conditions are met, it's a high card hand
  return("High Card")
}

# Function to simulate a round of poker for a specified number of players
deal_poker_round <- function(num_players) {
  if (num_players < 2 || num_players > 10) {
    cat("Number of players must be between 2 and 10.")
    return(NULL)
  }
  
  # Define the deck of cards
  suits <- c("Hearts", "Diamonds", "Clubs", "Spades")
  ranks <- c(2:10, "Jack", "Queen", "King", "Ace")
  deck <- expand.grid(rank = ranks, suit = suits)
  
  # Shuffle the deck
  #set.seed(42) # For reproducibility
  shuffled_deck <- deck[sample(nrow(deck)), ]
  
  # Deal cards to each player without replacement
  player_hands <- vector("list", num_players) #This is for initialization, the "player_hands" is an empty list, used to store the future results.
  for (i in 1:num_players) {
    # Randomly choose 5 unique cards from the shuffled deck
    player_indices <- sample(1:nrow(shuffled_deck), size = 5, replace = FALSE) 
    player_hand <- shuffled_deck[player_indices, ]
    
    # Remove the dealt cards from the shuffled deck (this line of code ensures no duplicate)
    shuffled_deck <- shuffled_deck[-player_indices, ]
    
    cat("Player", i, "Hand:")
    print(player_hand)
    hand_name <- get_poker_hand_name(player_hand$suit, player_hand$rank)
    cat("Hand Name:", hand_name, "\n\n")  #print out the results
    
    player_hands[[i]] <- player_hand
  }
}

# Example usage:
deal_poker_round(4)  # Simulate a round of poker with 4 players
```

(1) Yes, the inputs and outputs are as described.  
(2) Yes, they are valid cards with no duplicates.
(3) Yes, they are correct.
(4) Yes, the sampling part ensures that it is without replacement. If the number of player is more than 10, it gives "Number of players must be between 2 and 10." and then a Null set.